\section{Related Work}\label{sec:related-work}
\textbf{Sycophancy in Psychology.}
The study of sycophancy in psychology spans several research traditions, including impression management \cite{goffman1955facework}, social conformity \cite{asch1956studies}, ingratiation \cite{jones1964ingratiation}, and social influence \cite{cialdini1984influence, cialdini2016presuasion}, each capturing related but distinct aspects of the phenomenon. 
Conformity research has shown that individuals systematically adjust their judgements under group consensus, even when the group's position contradicts their private view \cite{asch1956studies, deutsch1955study}. 
A parallel line of work locates these adjustments to management of self presentation: Goffman \cite{goffman1955facework} characterized social interactions as dictated by face-work. Jones \cite{jones1964ingratiation} formalized normative influence as a reward strategy through the ingratiation framework, which consists of three primary tactics: other-enhancement, opinion conformity, and self-presentation. This framework has been extensively applied in the study of sycophancy in organizational settings \cite{higgins2003influence, higgins2004applicant, westphal2006boardroom}. Cialdini~\cite{cialdini1984influence} further distilled social influence research into six principles: reciprocity, commitment, social proof, liking, authority, and scarcity. More recently, empirical work has begun to operationalize perceived sycophancy, where \cite{rehani2026} developed a psychometrically validated scale revealing that it is captured by three distinct dimensions: uncritical agreement, obsequiousness, and excitement. Despite the richness of this field, existing sycophancy benchmarks have drawn on it only partially.

Research on social conformity has long demonstrated that individuals systematically adjust their judgments under group consensus \cite{asch1956studies}, driven by informational and normative influence \cite{deutsch1955study}. Two parallel 
theoretical traditions have elaborated on the mechanisms of normative influence. Goffman~\cite{goffman1955facework} established that social interactions are governed by face preservation, while Jones~\cite{jones1964ingratiation} formalized 
normative influence as a reward strategy through the ingratiation 
framework, which consists of three primary tactics: other-enhancement, opinion conformity, and self-presentation. This framework has been extensively applied in the study of sycophancy in organizational settings \cite{higgins2003influence, higgins2004applicant, westphal2006boardroom}. Cialdini~\cite{cialdini1984influence} 
further distilled social influence research into six principles: reciprocity, 
commitment, social proof, liking, authority, and scarcity. More recently, empirical work has begun to operationalize perceived sycophancy, where \cite{rehani2026} developed a psychometrically validated scale revealing that it is captured by three distinct dimensions: uncritical agreement, obsequiousness, and excitement. Despite the richness of  this field, existing sycophancy benchmarks have drawn on it only partially.

\textbf{Sycophancy in LLMs.}
Sycophancy, the tendency of a model to validate a user's stated beliefs or flatter 
Sycophantic behavior was first quantified systematically by \cite{perez2023modelwrittenevals}, who used model-written evaluations to show that LLMs trained with reinforcement learning from human feedback (RLHF) shift their stated opinions to align with user views. 
Subsequent mitigation work has proposed synthetic data augmentation~\cite{wei2023syntheticsycophancy}, direct preference optimization~\cite{khan2024mitigating}, pinpoint tuning~\cite{chen2024yes}, and probing \cite{papadatos2024probepenalty}, with~\cite{malmqvist2024sycophancymitigations} surveying their relative efficacy. 
On the evaluation side, \textsc{SycophancyEval}\cite{sharma2023towards} and \textsc{SycEval} \cite{SycEval} extended measurement to open-ended single turn tasks and the scope has since expanded to domain-specific settings such as mathematical theorem proving \cite{BROKENMATH}, medicine \cite{chen2025helpfulness}, and political opinion \cite{GermanPartiesQA}. In addition, in the single turn settings, there has been broader robustness testing \cite{PARROT, BEACON, DARKBENCH, duffy2024sycobench}. 
These benchmarks are predominantely single turn. Recent benchmarks such as \textsc{SYCON Bench} have examined the multi-turn setting \cite{TRUTHDECAY, FLIPFLOP, SYCONBENCH}. Beyond explicit change in opinion, \cite{ELEPHANT} introduce \textsc{ELEPHANT} which is grounded in social psychological theory to capture more implicit forms of sycophantic behavior. A complementary line of work examines how sycophancy manifests within model reasoning rather than final outputs, finding that social pressure 
induces post-hoc rationalization rather than faithful reasoning \cite{turpin2023languagemodels, turpin2023language, chang2026thermodynamic, walden2026reasoning, chua2025reasoning}, influences remain not verbalized and the correct reasoningcan preced sycophantic answers.  This work brings together the multi-turn pressure dimension, grounded social dimensions of sycophancy, and the role of user influence as a trigger into a unified evaluation framework.

\textbf{Mechanistic Understanding of LLM Sycophancy.}
Several studies have shown that sycophancy is encoded in linear directions within the activation space \cite{jain2025psychometrictraits, papadatos2024probepenalty, chen2025personavectors, rimsky2024steering, vennemeyer2025causalsycophancy, genadi2026linear}. 
However, much of this literature has implicitly treated sycophancy as a unitary construct \cite{rimsky2024steering, chen2025personavectors, papadatos2024probepenalty}. 
Recent findings argue against this assumption, with sparse autoencoder analyses \cite{templeton2024scaling} revealing non-overlapping residual-stream features associated with distinct sycophancy-related behaviors and causal steering interventions \cite{vennemeyer2025causalsycophancy} demonstrating that sycophantic agreement and praise occupy distinct subspaces that can be independently amplified or suppressed.
The conditions that trigger each subtype are less understood. As a first step, \cite{wang2025truthoverridden}, shows that different prompt formulations induce different degrees of sycophantic expression. Mechanistic work has also revealed that sycophantic commitment is detectable before thinking tokens are generated, but evolves throughout the reasoning trace\cite{feng2026good, mirtaheri2026catching, duszenko2026anchors, hu2026MONICA}. Together, these findings motivate an evaluation framework that goes beyond measuring sycophancy as a single score, instead systematically varying triggers to elicit distinct subtypes of sycophantic behavior.


\input{tables/BenchScope}

